---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Reading The Data In
#### After downloading the file from [https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"]

#### Read in the data set.
```{r}
fitnessData <- read.csv("~/Coursera_R/fitbit/activity.csv") 
```

## A histogram of the total number of steps taken each day.

```{r}
# First need to aggeragate data by day.
dailyStepsSumByDay <- aggregate(fitnessData$steps, by=list(fitnessData$date), FUN=sum,na.rm=TRUE) 
# Plot historgram of tiotla number of steps taken each day.
hist(dailySteps$x)
```

#### Then calculate mean by day:
```{r}
mean(dailySteps$x,)
```
#### Then calculate median by day:
```{r}
median(dailySteps$x,)
```

## What is the average daily activity pattern?
####Determine the daily activity pattern by aggregating by the 5 min. interval
```{r}
averageStepsPerInterval <-aggregate(fitnessData$steps, by=list(fitnessData$interval), FUN=mean,na.rm=TRUE) 
```
#### Plot the average number of steps taken (averaged across all days) versus the 5-minute intervals.
```{r}
plot(averageStepsPerInterval,type="l", xlab= "5 minute Interval", ylab= "Average Steps") 
```

## Report the 5-minute interval that, on average, contains the maximum number of steps.
```{r}
averageStepsPerInterval[which.max(averageStepsPerInterval$x),]
```


## Imputing missing values
#### Impute the missing base NA's using the Interval average.
#### Merge the intervalaverage data set and base data set
```{r}
mergedFitnessData <- merge(fitnessData,averageStepsPerInterval, by.x = "interval", by.y="Group.1",all.x=TRUE) 
```
#### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imputedFitnessData <- transform(mergedFitnessData, imputedSteps = ifelse(is.na(steps), x, steps)) 
```

## A histogram of the total number of steps taken each day after missing values were imputed.
```{r}
#Aggergate by day
dailyImputedStepsSumByDay <- aggregate(fitnessData$steps, by=list(imputedFitnessData$date), FUN=sum,na.rm=TRUE) 

# Plot historgram of imputed steps by day
hist(imputedFitnessData$imputedSteps)
```

## Are there differences in activity patterns between weekdays and weekends?  The grpahs below show greater activity across intervals on the weekends.
```{r}
# Determine Day of Week
imputedFitnessData$dayofweek <- weekdays(as.Date(imputedFitnessData$date))

# Determine Weekends
weekendImputedDate <- transform(imputedFitnessData, dayType = ifelse(imputedFitnessData$dayofweek=="Saturday"|imputedFitnessData$dayofweek=="Sunday","Weekend","Weekday")) 

averageImputedStepsPerInterval <-aggregate(imputedSteps ~ interval + dayType,data=weekendImputedDate,  FUN=mean) 

#plot the average number of steps taken per 5-minute interval across weekdays and weekends

library(lattice)
head(averageImputedStepsPerInterval)
xyplot(imputedSteps ~ interval| dayType, data=averageImputedStepsPerInterval, type="l", xlab = "5 min. interval", ylab = "steps", main = "Weekday vs Weekend steps per 5 min. Interval")
```


